The performance of presence-based and process-based species distribution models under realistic conditions

### 2020-02-13 ms/ms_latex/ms_sdmComp.tex
Updates from Overleaf

\section{Results}
\label{S:3}
% Talk results
% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
% 2. Under ideal conditions, process-based outperform Maxent
% 3. For non-equilibrium distributions, process-based outperform Maxent
% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
% 5. Dispersal is critical for CA models
% 6. Structural errors seem to have erratic effects
% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty

@@

% Effect of modelling issues
modelling with \emph{incorrect covariates} was universally problematic for model performance, though it had a greater impact on TSS in the shrub than the biennial species (Fig. \ref{fig:TSSvIdeal}). \emph{Incorrect covariates} did not lead to systematic over-prediction or under-prediction, but rather a decrease in both sensitivity and specificity, particularly for the process-based models (Fig. B.3, Fig. B.4). Excluding the seedbank in the process-based models had minimal effect on TSS for the shrub, but variable effects on TSS for the biennial. Relative to \emph{ideal}, all process-based models showed reduced sensitivity and increased specificity for the biennial, indicating a smaller predicted range. The IPM was generally more resistant to mis-characterized dispersal compared to the simulation-based CA\textsubscript{i} and CA\textsubscript{p} models. Underestimation of dispersal parameters led to decreased predicted ranges, and overestimation to increased predicted ranges relative to ‘ideal’. The effects were more extreme for the biennial than for the shrub.

@@

Similarly, the CA models predict the abundance $N$ for each species. Discrepancies in predicted abundance from the CA\textsubscript{p} model showed a spatial banding of upward bias at the range margins for both species (Fig. \ref{fig:N_Lambda}), indicating a tendency to over-predict at range edges. Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} model showed upward bias for the shrub (Fig. \ref{fig:N_Lambda}). The CA\textsubscript{p} model dramatically underestimated abundance of the long-lived shrub with \emph{non-equilibrium} datasets, despite minimal impact on predicted presence or absence. As with the IPM, using \emph{incorrect covariates} tended to result in larger error in abundance predictions for both CA\textsubscript{p} and CA\textsubscript{i}.

@@

% Talk Discussion
% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
% 2. Under ideal conditions, process-based outperform Maxent
% 3. For non-equilibrium distributions, process-based outperform Maxent
% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
% 5. Dispersal was critical for CA models, but had essentially no effect for IPMs
% 6. Structural errors seem to have erratic effects
% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty

@@

% Dangers of process-based SDMs: dispersal
The accuracy of the CA model predictions were sensitive to dispersal. Drastically underestimating dispersal resulted in smaller predicted ranges, while overpredicting dispersal resulted in larger predicted ranges. The predictions for the biennial tended to be a bit more sensitive, particularly in terms of specificity (Fig. B.4).

@@

@@

\section{Bibliography}
\bibliography{sdmMethodComp}

>>>>

\section{Results}
\label{S:3}

@@

% Effect of modelling issues
Modelling with \emph{incorrect covariates} decreased performance for all SDMs, though the decline tended to be larger on the process-based SDMs. It generally had a greater impact on TSS in the shrub than the biennial species (Fig. \ref{fig:TSSvIdeal}). \emph{Incorrect covariates} did not lead to systematic over-prediction or under-prediction, but rather a decrease in both sensitivity and specificity, particularly for the process-based models (Fig. B.3, Fig. B.4). Excluding the seedbank in the process-based models had minimal effect on TSS for the shrub, but variable effects on TSS for the biennial. Relative to \emph{ideal}, all process-based models showed reduced sensitivity and increased specificity for the biennial, indicating a smaller predicted range. The IPM was generally more resistant to mis-characterized dispersal compared to the simulation-based CA\textsubscript{i} and CA\textsubscript{p} models. Underestimation of dispersal parameters led to decreased predicted ranges, and overestimation to increased predicted ranges relative to \emph{ideal}. The effects were more extreme for the biennial than for the shrub.

@@

Similarly, the CA models predict the abundance $N$ for each species. Discrepancies in predicted abundance from the CA\textsubscript{p} model showed a spatial banding of upward bias at the range margins for both species (Fig. \ref{fig:N_Lambda}), indicating a tendency to over-predict at range edges. Immediately interior to the range margin, the CA\textsubscript{p} and CA\textsubscript{i} models showed upward bias for the shrub (Fig. \ref{fig:N_Lambda}). The CA\textsubscript{p} model dramatically underestimated abundance of the long-lived shrub with \emph{non-equilibrium} datasets, despite minimal impact on predicted presence or absence. As with the IPM, using \emph{incorrect covariates} tended to result in larger error in abundance predictions for both CA\textsubscript{p} and CA\textsubscript{i}.

@@

@@

% Dangers of process-based SDMs: dispersal
The accuracy of the CA model predictions were sensitive to dispersal (Fig. \ref{fig:TSSvIdeal}). As expected, underestimating dispersal resulted in smaller predicted ranges, while overestimating dispersal resulted in larger predicted ranges. The predictions for the biennial tended to be a bit more sensitive, particularly in terms of specificity (Fig. B.4). In contrast, the IPM predictions were less affected by errors in the dispersal parameters (Fig. \ref{fig:TSSvIdeal}).

@@

\section{Bibliography}
\bibliography{ms/ms_latex/sdmMethodComp}

%% RESULTS
% Talk results
% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
% 2. Under ideal conditions, process-based outperform Maxent
% 3. For non-equilibrium distributions, process-based outperform Maxent
% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
% 5. Dispersal is critical for CA models
% 6. Structural errors seem to have erratic effects
% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty

%% DISCUSSION
% Talk Discussion
% 1. Across all scenarios, IPM is best for lambda, CAi/p are best for N, with Maxent in the middle for both
% 2. Under ideal conditions, process-based outperform Maxent
% 3. For non-equilibrium distributions, process-based outperform Maxent
% 4. Process-based are much, much more sensitive to covariate choice -- implies sensitivity to regression accuracy
% 5. Dispersal was critical for CA models, but had essentially no effect for IPMs
% 6. Structural errors seem to have erratic effects
% 0. Process-based are better IF you are confident of the process, but Maxent is more robust to error and uncertainty

@@
